Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-09-04 paper.Rmd
add appendix figure with diff't smoothing methods
extends Figure 2 that shows the benefit of (binormal) smoothing

>>>>

## Different smoothing methods

The **pROC** package `smooth()` function includes several different smoothing methods. The default, which BS use in their code, is binormal smoothing. Here is a replication of the smooth benefits shown in Figure \ref{fig:benefit-plot} with two other smoothing methods. (We could not get "logcondens" and "logcondens.smooth" to work, and they are based on another external package that has not been updated since 2016.)

```{r smooth-benefit-plot-extended, fig.height=3.5, fig.width=5.5, out.width = ".8\\linewidth", fig.align = "center", fig.cap = "Gain from using smoothed ROC to calculate AUC, for each model reported in Tables 3 and 4 (B\\&S Tables 1 and 2), with 3 different smoothing methods (binormal is the default method and used in B\\&S)"}

suppressPackageStartupMessages({
  library(tidyverse)
  library(yardstick)
  library(pROC)
  library(forcats)
})

preds <- read_rds("rep_nosmooth/output/all-predictions.rds")

auc_roc_vec <- function(pred, truth, smooth, sm = "binormal") {
  roc_obj <- pROC::roc(truth, pred, auc = TRUE, quiet = TRUE, smooth = smooth, smooth.method = sm)
  as.numeric(roc_obj$auc)
}

rocs <- preds %>% 
  filter(!is.na(value)) %>%
  group_by(cell_id) %>% 
  summarize(
    no = auc_roc_vec(pred = pred, truth = value, smooth = FALSE),
    binormal = auc_roc_vec(pred = pred, truth = value, smooth = TRUE, sm = "binormal"),
    density = auc_roc_vec(pred = pred, truth = value, smooth = TRUE, sm = "density"),
    fitdistr = auc_roc_vec(pred = pred, truth = value, smooth = TRUE, sm = "fitdistr")
  )

# these two take a long time and error out for some reason
#logcondens = auc_roc_vec(pred = pred, truth = value, smooth = TRUE, sm = "logcondens"),
#logcondens.smooth = auc_roc_vec(pred = pred, truth = value, smooth = TRUE, sm = "logcondens.smooth")

mt <- read_rds("rep_nosmooth/output/model-table-w-results.rds")
mt <- mt %>% left_join(rocs, by = "cell_id")

model_levels <- c("Escalation", "Quad", "Goldstein", "CAMEO", "Average", 
                  "With PITF Predictors", "Weighted by PITF", "PITF Split Population",
                  "PITF Only")

benefit <- mt %>% 
  mutate(column = fct_recode(column, Escalation = "Escalation Only"),
         column = factor(column, levels = rev(model_levels))) %>%
  mutate(binormal = binormal - no,
         density  = density - no,
         fitdistr = fitdistr - no) %>%
  select(cell_id:non_RF, binormal:fitdistr) %>%
  pivot_longer(binormal:fitdistr, names_to = "method")

means <- benefit %>%
  group_by(column, method) %>%
  summarize(avg_benefit = mean(value))

ggplot(benefit, aes(x = value, y = column, color = column)) +
  facet_wrap(~method) +
  geom_point() +
  scale_color_manual(guide = FALSE, values = cols) +
  theme(axis.text.y = element_text(size = 12)) +
  geom_point(data = means, aes(x = avg_benefit), shape = "|", color = "black", size = 6) +
  labs(x = "AUC(smoothed ROC) - AUC(empirical ROC)",
       y = "") +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 3) +
  theme(axis.text.y = element_text(size = 12))
```
